Supplemental Information

### 2021-12-20 manuscript/manuscript.Rmd
reformat for nature comms and fix figure legend (*) by escaping

bibliography: 
- library.bib
- packages.bib
csl: cell-metabolism.csl
link-citations: yes
compact-title: true
geometry: margin = 0.75in
fontsize: 11pt
mainfont: Calibri
graphics: yes
---

@@

# Summary

@@

(ref:m1) **Effects of prolyl hydroxylase inhibition on extracellular metabolite fluxes in lung fibroblasts.** Lung fibroblasts (LFs) were cultured with 21% oxygen (*red*), 0.5% oxygen (*blue*), DMSO (0.1% v/v, *green*), or molidustat (BAY, 10 μM, *purple*) beginning 24 h prior to time 0. (**A**) Growth curves of LFs in each experimental condition (n = 8). (**B**) Growth rates from (A) were determined by robust linear modeling of log-transformed growth curves. (**C**, **D**) Representative immunoblots of LF protein lysates cultured as in (A). (**E**, **F**) Relative change in HIF-1α (E) and LDHA (F) protein levels normalized to 21% oxygen or DMSO treatment time 0 (n = 4). (**G**, **H**) Relative changes in GLUT1 (G) and LDHA (H) mRNA levels normalized to 21% oxygen or DMSO treatment time 0 (n = 4). (**I**, **J**) Extracellular fluxes of the indicated metabolites (n = 8). Data are mean ± SEM. Comparisons were made using linear mixed effects models with treatment group as a fixed effect and biological replicate as a random effect. Tukey's *post hoc* test was applied to determine differences between 21% and 0.5% oxygen (*), between DMSO and BAY treatment (†), or between 0.5% oxygen and BAY treatment (‡) with adjusted p-values < 0.05 considered significant.

(ref:m2) **Stable isotope tracing of lung fibroblasts.** (**A**) Fraction of pyruvate labeling following treatment of lung fibroblasts (LFs) with [U-^13^C~6~] glucose. (**B**) Rate values determined from asymptotic regression fit of the data from (A). (**C**) Isotopic labeling of key intracellular metabolites after 72 h of treatment with the indicated tracers. Experimentally determined mass isotope distributions were corrected for natural isotope abundance. Data are the mean ± SEM of 4 biological replicates. Significant differences in labeling patterns between 21% and 0.5% oxygen (*), DMSO and BAY treatment (†), and 0.5% oxygen and BAY treatment (‡) for each combination of metabolite and tracer are highlighted. FBP, fructose bisphosphate; 3PG, 3-phosphoglycerate; AKG, α-ketoglutarate.

@@

(ref:m4) **PHD inhibition increases lactate uptake and oxidation.** Mass isotopomer distributions of key metabolites following labeling with [U-^13^C~3~] lactate (2 mM) for 72 h indicates increased lactate uptake and oxidation in hypoxia or with molidustat (BAY) treatment. Data are mean ± SEM of n = 4 biological replicates. Comparisons were made using linear mixed effects models with treatment group as a fixed effect and biological replicate as a random effect. Tukey's *post hoc* test was applied to determine differences between 21% and 0.5% oxygen (*), between DMSO and BAY treatment (†), or between 0.5% oxygen and BAY treatment (‡) with adjusted p-values < 0.05 considered significant.

@@

# STAR Methods

@@

### Lung fibrolbasts

>>>>

bibliography: 
- library.bib
- packages.bib
csl: nature-communications.csl
link-citations: yes
compact-title: true
geometry: margin = 0.75in
fontsize: 11pt
mainfont: Calibri
graphics: yes
---

@@

# Abstract

@@

(ref:m1) **Effects of prolyl hydroxylase inhibition on extracellular metabolite fluxes in lung fibroblasts.** Lung fibroblasts (LFs) were cultured with 21% oxygen (*red*), 0.5% oxygen (*blue*), DMSO (0.1% v/v, *green*), or molidustat (BAY, 10 μM, *purple*) beginning 24 h prior to time 0. (**A**) Growth curves of LFs in each experimental condition (n = 8). (**B**) Growth rates from (A) were determined by robust linear modeling of log-transformed growth curves. (**C**, **D**) Representative immunoblots of LF protein lysates cultured as in (A). (**E**, **F**) Relative change in HIF-1α (E) and LDHA (F) protein levels normalized to 21% oxygen or DMSO treatment time 0 (n = 4). (**G**, **H**) Relative changes in GLUT1 (G) and LDHA (H) mRNA levels normalized to 21% oxygen or DMSO treatment time 0 (n = 4). (**I**, **J**) Extracellular fluxes of the indicated metabolites (n = 8). Data are mean ± SEM. Comparisons were made using linear mixed effects models with treatment group as a fixed effect and biological replicate as a random effect. Tukey's *post hoc* test was applied to determine differences between 21% and 0.5% oxygen (\*), between DMSO and BAY treatment (†), or between 0.5% oxygen and BAY treatment (‡) with adjusted p-values < 0.05 considered significant.

(ref:m2) **Stable isotope tracing of lung fibroblasts.** (**A**) Fraction of pyruvate labeling following treatment of lung fibroblasts (LFs) with [U-^13^C~6~] glucose. (**B**) Rate values determined from asymptotic regression fit of the data from (A). (**C**) Isotopic labeling of key intracellular metabolites after 72 h of treatment with the indicated tracers. Experimentally determined mass isotope distributions were corrected for natural isotope abundance. Data are the mean ± SEM of 4 biological replicates. Significant differences in labeling patterns between 21% and 0.5% oxygen (\*), DMSO and BAY treatment (†), and 0.5% oxygen and BAY treatment (‡) for each combination of metabolite and tracer are highlighted. FBP, fructose bisphosphate; 3PG, 3-phosphoglycerate; AKG, α-ketoglutarate.

@@

(ref:m4) **PHD inhibition increases lactate uptake and oxidation.** Mass isotopomer distributions of key metabolites following labeling with [U-^13^C~3~] lactate (2 mM) for 72 h indicates increased lactate uptake and oxidation in hypoxia or with molidustat (BAY) treatment. Data are mean ± SEM of n = 4 biological replicates. Comparisons were made using linear mixed effects models with treatment group as a fixed effect and biological replicate as a random effect. Tukey's *post hoc* test was applied to determine differences between 21% and 0.5% oxygen (\*), between DMSO and BAY treatment (†), or between 0.5% oxygen and BAY treatment (‡) with adjusted p-values < 0.05 considered significant.

@@

# Methods

@@

### Lung fibroblasts

### 2021-12-20 manuscript/supplement.Rmd
reformat for nature comms and fix figure legend (*) by escaping

bibliography_main: library.bib
bibliography_software: packages.bib
csl: cell-metabolism.csl
link-citations: yes
compact-title: true
geometry: margin = 0.75in
fontsize: 11pt
mainfont: Calibri
graphics: yes

@@

(ref:s2) **Extracellular flux measurements in 0.2% oxygen.** Lung fibroblasts (LFs) were cultured with 21% oxygen (*red*) or 0.2% oxygen (*dark blue*) beginning 24 h prior to time 0. (**A**) Growth curves of LFs in each experimental condition (n = 4). (**B**) Growth rates from (A) were determined by robust linear modeling of log-transformed growth curves. (**C**) Representative immunoblot of LF protein lysates cultured as in (A). (**D**, **E**) Relative change in HIF-1α (D) and LDHA (E) protein levels normalized to 21% oxygen time 0 (n = 4). (**F**, **G**) Relative changes in GLUT1 (G) and LDHA (H) mRNA levels normalized to 21% oxygen time 0 (n = 4). (**H**, **I**) Extracellular fluxes of the indicated metabolites (n = 4). Data are mean ± SEM. Comparisons were made using linear mixed effects models with treatment group as a fixed effect and biological replicate as a random effect. Tukey's *post hoc* test was applied to determine differences between 21% and 0.2% oxygen (*) with p-values < 0.05 considered significant.

(ref:s3) **Extracellular flux measurements in pulmonary artery smooth muscle cells in 0.5% oxygen.** Pulmonary artery smooth muscle cells (PASMCs) were cultured with 21% oxygen (*red*) or 0.5% oxygen (*blue*) beginning 24 h prior to time 0. (**A**) Growth curves of PASMCs under in each experimental condition (n = 8). (**B**) Growth rates from (A) were determined by robust linear modeling of log-transformed growth curves. (**C**) Representative immunoblot of PASMC protein lysates cultured as in (A). (**D**, **E**) Relative change in HIF-1α (D) and LDHA (E) protein levels normalized to 21% oxygen time 0 (n = 4). (**F**, **G**) Relative changes in GLUT1 (G) and LDHA (H) mRNA levels normalized to 21% oxygen time 0 (n = 4). (**H**, **I**) Extracellular fluxes of the indicated metabolites (n = 8). Data are mean ± SEM. Comparisons were made using linear mixed effects models with treatment group as a fixed effect and biological replicate as a random effect. Tukey's *post hoc* test was applied to determine differences between 21% and 0.5% oxygen (*) with p-values < 0.05 considered significant.

(ref:s4) **Mass isotopomer distributions after 72 h of labeling in lung fibroblasts.** Lung fibroblasts (LFs) were labeled with the indicated tracers and intracellular metabolites were analyzed by LC-MS after 72 h. Mass isotopomer distributions were adjusted for natural abundance. Data are the mean ± SEM of 4 biological replicates. Significant differences in labeling patterns between 21% and 0.5% oxygen (*), DMSO and BAY treatment (†), and 0.5% oxygen and BAY treatment (‡) for each combination of metabolite and tracer are highlighted.

(ref:s5) **Mass isotopomer distributions after 72 h of labeling in pulmonary artery smooth muscle cells.** Pulmonary artery smooth muscle cells (PASMCs) were labeled with the indicated tracers and intracellular metabolites were analyzed by LC-MS after 36 h. Mass isotopomer distributions were adjusted for natural abundance. Data are the mean ± SEM of 4 biological replicates. Significant differences in labeling patterns between 21% and 0.5% oxygen (*) for each combination of metabolite and tracer are highlighted.

>>>>

bibliography_main: library.bib
bibliography_software: packages.bib
csl: nature-communications.csl
link-citations: yes
compact-title: true
geometry: margin = 0.75in
fontsize: 11pt
mainfont: Calibri
graphics: yes

@@

(ref:s2) **Extracellular flux measurements in 0.2% oxygen.** Lung fibroblasts (LFs) were cultured with 21% oxygen (*red*) or 0.2% oxygen (*dark blue*) beginning 24 h prior to time 0. (**A**) Growth curves of LFs in each experimental condition (n = 4). (**B**) Growth rates from (A) were determined by robust linear modeling of log-transformed growth curves. (**C**) Representative immunoblot of LF protein lysates cultured as in (A). (**D**, **E**) Relative change in HIF-1α (D) and LDHA (E) protein levels normalized to 21% oxygen time 0 (n = 4). (**F**, **G**) Relative changes in GLUT1 (G) and LDHA (H) mRNA levels normalized to 21% oxygen time 0 (n = 4). (**H**, **I**) Extracellular fluxes of the indicated metabolites (n = 4). Data are mean ± SEM. Comparisons were made using linear mixed effects models with treatment group as a fixed effect and biological replicate as a random effect. Tukey's *post hoc* test was applied to determine differences between 21% and 0.2% oxygen (\*) with p-values < 0.05 considered significant.

(ref:s3) **Extracellular flux measurements in pulmonary artery smooth muscle cells in 0.5% oxygen.** Pulmonary artery smooth muscle cells (PASMCs) were cultured with 21% oxygen (*red*) or 0.5% oxygen (*blue*) beginning 24 h prior to time 0. (**A**) Growth curves of PASMCs under in each experimental condition (n = 8). (**B**) Growth rates from (A) were determined by robust linear modeling of log-transformed growth curves. (**C**) Representative immunoblot of PASMC protein lysates cultured as in (A). (**D**, **E**) Relative change in HIF-1α (D) and LDHA (E) protein levels normalized to 21% oxygen time 0 (n = 4). (**F**, **G**) Relative changes in GLUT1 (G) and LDHA (H) mRNA levels normalized to 21% oxygen time 0 (n = 4). (**H**, **I**) Extracellular fluxes of the indicated metabolites (n = 8). Data are mean ± SEM. Comparisons were made using linear mixed effects models with treatment group as a fixed effect and biological replicate as a random effect. Tukey's *post hoc* test was applied to determine differences between 21% and 0.5% oxygen (\*) with p-values < 0.05 considered significant.

(ref:s4) **Mass isotopomer distributions after 72 h of labeling in lung fibroblasts.** Lung fibroblasts (LFs) were labeled with the indicated tracers and intracellular metabolites were analyzed by LC-MS after 72 h. Mass isotopomer distributions were adjusted for natural abundance. Data are the mean ± SEM of 4 biological replicates. Significant differences in labeling patterns between 21% and 0.5% oxygen (\*), DMSO and BAY treatment (†), and 0.5% oxygen and BAY treatment (‡) for each combination of metabolite and tracer are highlighted.

(ref:s5) **Mass isotopomer distributions after 72 h of labeling in pulmonary artery smooth muscle cells.** Pulmonary artery smooth muscle cells (PASMCs) were labeled with the indicated tracers and intracellular metabolites were analyzed by LC-MS after 36 h. Mass isotopomer distributions were adjusted for natural abundance. Data are the mean ± SEM of 4 biological replicates. Significant differences in labeling patterns between 21% and 0.5% oxygen (\*) for each combination of metabolite and tracer are highlighted.
